### 2023-03-23 reachable-resilience.tex
Changed name of Uniform resilience to remove 'Uniform'
up

\section{Resilience for WSTS}

Uniform resilience is a strong property since it implies that from bad there must exist a path to Safe. But, when one considers a system with an initial state $s_0$, it could be sufficient to ask that only from $post^*(s_0) \cap Bad$, there must exist a path to Safe. However, this condition seems more difficult since one also must decide whether a state $s \in Bad$ is reachable from $s_0$.
%
%\problemx{(I,J)-$k$-resilience problem for WSTS}
%{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis), a set $J$.}
%{$\forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
%
%\problemx{general $k$-resilience problem for WSTS}
%{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis).}
%{$\forall s'  ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}

@@

\problemx{$k$-resilience problem for WSTS}
{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $U$ with a given basis, a decidable downward-closed set $D$.}
{$D \longrightarrow^{\leq k} U$ ?  ou $D \cap post^*(s) \longrightarrow^{\leq k} U$ ?  \newline}
%
%  		{$\forall s' ~ (s \rightarrow^* s') \wedge s' \in D  \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\newline}
%
$D \longrightarrow^{\leq k} U$: $\forall s' \in D \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?\\
%
$D \cap post^*(s) \longrightarrow^{\leq k} U$ ?
%
$\forall s' ~ (s \rightarrow^* s') \wedge s' \in D  \implies \exists s'' \in U ~ s' \rightarrow^{\leq k} s''$ ?
 \alain{ça signifie pour tout $d \in D$ il existe un chemin de longueur $\leq k$ qui arrive dans un $u \in U$. c'est plus simple et plus général et ça ne demande moins de décider l'accessibilité....comment faire pour vérifier pour tous les (D,U) ? sinon ça dépend tjs de l'état intial $s$ et s'il faut vérifier pour tous les $s$....}

%ou plus simple : $D \rightarrow^{\leq k} U$ ?

>>>>

\section{Resilience from a single-state for WSTS}

Resilience is a strong property since it implies that from every element of $\Bad$ there must exist a path to $\Safe$. But, when one considers a system with an initial state $s_0$, it could be sufficient to ask that only from $post^*(s_0) \cap \Bad$, there must exist a path to $\Safe$. However, this condition seems more difficult since one also must decide whether a state $s \in \Bad$ is reachable from $s_0$.
%
%\problemx{(I,J)-$k$-resilience problem for WSTS}
%{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis), a set $J$.}
%{$\forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
%
%\problemx{general $k$-resilience problem for WSTS}
%{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis).}
%{$\forall s'  ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}

@@

\problemx{$k$-resilience problem for WSTS}
{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an upward-closed set $\Safe$ with a given basis, a decidable downward-closed set $\Bad$.}
{ $\Bad \cap \post^*(s) \longrightarrow^{\leq k} \Safe$ ?  \newline}
%

### 2023-03-23 resilience.tex
up

% \textcolor{red}{Ask the question why use a set of propositions for $\Safe$ and $\Bad$ rather than use subsets of the set of configurations ? }

@@

\begin{proposition}\label{general}
\textcolor{red}{
$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
(\Bad,\Safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
(\Bad,\Safe)-resilient}
\end{proposition}

@@

\begin{theorem}\label{k-down-up}
Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and the predbasis hypothesis, and two finite sets: $I_{\Bad}$ and $Min_{\Safe}$.
The  $k$-resilience problem (kRP) is decidable.
 \end{theorem}

@@

\begin{theorem}\label{down-down}
The resilience problem is undecidable for WSTS with \textcolor{red}{more than one minimal element}, $\Safe=\downarrow \Safe$
and $\Bad=\downarrow \Bad$.
\end{theorem}

\begin{proof}
In the case of a WSTS with more than one minimal element $0_1$ and $0_2$, then the problem of whether $0_2$ is reachable from $0_1$ reduces itself to the resilience problem.  
\end{proof}

In the case of a WSTS with only one minimal element,  
% (ex: Petri Nets and their «Zero») 
then by downward closure the minimal element is present in both \Safe and \Bad and hence resilience always hold.

@@

In a reset petri-net, if $\Bad = \uparrow \mu$ for some $\mu \neq 0$,
and $\Safe = \downarrow 0$,
then the resilience problem is basically \textcolor{red}{(is it?)} the single-state zero-reachability problem, and is undecidable.

\mathieu{So, the idea is that the question "is $0$ reachable from token assigment $\mu$" is undecidable, but here the question is slightly different because we start from $\uparrow \mu$ instead.}

\begin{center}
\begin{tabular}{ | l | c | c | c | r |}
\hline   \Safe~\Bad & $\uparrow$~ $\uparrow$~ & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~ \\ \hline
   RP & Decidable (Theorem~\ref{up-up}) & Decidable (Theorem~\ref{down-up}) & Undecidable (Theorem~\ref{down-down}) & \textcolor{red}{Undecidable} \\ \hline
   BRP & Decidable (Corollary~\ref{B-up-up}) &  Decidable (Theorem~\ref{down-up}) & Undecidable & ?? \\ \hline
      kRP & Decidable (Theorem~\ref{k-up-up}) & \textcolor{red}{Decidable} (Theorem~\ref{k-down-up}) & Undecidable & ?? \\ \hline
 \end{tabular}
\end{center}

@@

>>>>

@@

\begin{proposition}\label{general}
\textcolor{red}{
$\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
(\Bad,\Safe)-bounded-resilient iff $\mathscr{S}=(S,\rightarrow,\leq)$ is %uniformely 
(\Bad,\Safe)-resilient (CONJECTURE FOR NOW)
}
\end{proposition}

@@

\begin{theorem}\label{k-down-up}
Let $\mathscr{S}=(S,\rightarrow, \leq)$ be a completion-post-effective $\omega^2$-WSTS with strong compatibility and the predbasis hypothesis, and two finite sets: $I_{\Bad}$ and $Min_{\Safe}$.
The  $k$-resilience problem (kRP) is decidable. \textcolor{red}{CONJECTURE for now}
 \end{theorem}

@@

\begin{theorem}\label{down-down}
The resilience problem is undecidable for WSTS with more than one minimal element, 
$\Safe=\downarrow \Safe$
and $\Bad=\downarrow \Bad$.
\end{theorem}

\begin{proof}
In the case of a WSTS with at least two minimal elements $0_1$ and $0_2$, the problem of whether $0_2$ is reachable from $0_1$ reduces itself to the resilience problem, hence its undecidability.  
\end{proof}

In the case of a WSTS with only one minimal element,  
% (ex: Petri Nets and their «Zero») 
then by downward closure the minimal element is present in both $\Safe$ and $\Bad$ and hence resilience always hold.

@@

In a reset petri-net, if $\Bad = \uparrow \mu$ for some $\mu \neq 0$,
and $\Safe = \downarrow 0$,
then the resilience problem is similar to the single-state zero-reachability problem, which is undecidable.

So, the idea is that the question "is $0$ reachable from token assigment $\mu$" is undecidable, but here the question is slightly different because we start from $\uparrow \mu$ instead.

\subsection{Case Synthesis}

@@

\begin{center}
\begin{tabular}{ | l | c | c | c | r |}
\hline   \Safe~\Bad & $\uparrow$~ $\uparrow$~ & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~ \\ \hline
   RP & Decidable (Theorem~\ref{up-up}) & Decidable (Theorem~\ref{down-up}) & Undecidable (Theorem~\ref{down-down}) & \textcolor{red}{Undecidable} \\ \hline
   BRP & Decidable (Corollary~\ref{B-up-up}) &  Decidable (Theorem~\ref{down-up}) & ?? & ?? \\ \hline
      kRP & Decidable (Theorem~\ref{k-up-up}) & \textcolor{red}{Decidable} (Theorem~\ref{k-down-up}) & ?? & ?? \\ \hline
 \end{tabular}
\end{center}
